---
title: Kinds of datasets and their licenses in Missouri's data portal
twitter_title: Missouri's data portal has lots of traffic surveys
tags: ['open-data']
---
<% root = 'https://github.com/tlevine/socrata-analysis/tree/master/' %>
```{r configure, echo = F}
opts_chunk$set(fig.width=10, fig.height=5, dpi = 42 * 5)
```

## Which specific datasets have licenses?
I recently [found](/!/open-data-licensing) that most datasets on the
open data portals I've been looking at don't have open licenses.
But a lot of datasets do have licenses, and licensing does vary within
a portal. So I wanted to see which datasets get licenses and which ones
don't.

## The Missouri Data Portal
I went about this by looking at one data portal,
[data.mo.gov](https://data.mo.gov). Here's why I chose it.

1. To make it easy to download the license information, I wanted to use
    Socrata's DCAT endpoint. This endpoint only returns up to 1000
    datasets, so I needed a portal with no more than 1000 datasets.
    Missouri's portal has `r nrow(missouri)` datasets.
2. I think I've figured out how to
    [deduplicate federated datasets](/!/),
    but I didn't want to deal with that.
    The Missouri portal federates only the
    [Kansas City data portal](https://data.kcmo.org).
3. Missouri has [a lot of PDF files](/!/),
    so I thought there might be something in there that would be more
    original than mere facts and covered under copyright; this would
    make a license more important.

I wound up categorizing the datasets quite a bit more than I originally
intended, so now I know a lot about the sorts of data on the portal.

## Licenses
Most datasets on the missouri portal are public domain.

```{r publicdomain}
print(p0)
```

Proportionately more PDF files are public domain.

```{r publicdomain.pdf}
print(p1)
```

And most PDFs are traffic surveys.

```{r pdf.trafficsurvey}
print(p2)
```

Also, public domain is the only license; if a dataset isn't licensed
public domain, it isn't licensed at all.

```{r na.table}
table(missouri$public.domain, is.na(missouri$license))
```

## Dataset kinds

## Future study
In the Missouri portal, I used only title to group datasets; I didn't
look at the long-form description, the author or the schema. There were
pretty clear-cut divisions between titles, but I would like to see how
reasonable this approach really is; I want to whether groups of similarly
named datasets tend to have similar schemas.
